---
title: "Figures for manuscript"
author: "Jonathan Harrison"
date: "06/11/2018"
output: html_document
fig_width: 6
fig_height: 4
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4,
                      echo=FALSE, warning=FALSE, message=FALSE)
```
Outline of Figures:

Fig1: Drosophila egg chambers exhibit a characteristic network structure of connections between nurse cells A) Cartoon of tissue, B) Equivalent image of tissue, C) Diagram of connections between NCs via RCs.


Fig2: Behaviour of differential equation model. A) Dynamics in time shown for each cell, B) Model related to network structure of nurse cells.
```{r, Fig2, echo=FALSE,message=FALSE}
source('simulate_from_ODE_model.R')
```

Fig3: Assembly of higher order complexes in oocyte A) Particles in different regions of the egg chamber, B) Distribution of intensities of particles in different regions.

```{r, Fig3, echo=FALSE,message=FALSE}
source('estimate_phi_directly.R')
b %>% 
  mutate(Normalised_intensity = BgdSubtract / median(a$MeanByRegion)) %>%
  ggplot(aes(x=Region,y=Normalised_intensity)) +
  geom_violin(draw_quantiles = c(0.5)) +
  geom_jitter(alpha=0.3) +
  labs(y='Normalised intensity') +
  scale_x_discrete("Region", labels = c("background" = "BGD","oocyte" = "OO","nurse_cells" = "NC")) +
  theme_bw()
ggsave('plots/fig3d.eps',device=cairo_ps)
```


Fig4: Characterisation of the bias in directionality of transport through ring canals A) Schematic of transport between two compartments, B) Posterior pairs plot, C) Posterior predictive plot.
![optional caption text](plots/fig4_schematic.eps){ width=30% }
```{r, Fig4, echo=FALSE,message=FALSE}
source('run_model_comparison_stst.R')
identifier = 'MCv500'
res <- run_model_comparison_stst(identifier=identifier,
                                 nSamples=9,nTest=0,nTestOE=0,
                                 use_real_data=TRUE,run_mcmc=FALSE,
                                 parametersToPlot = c('nu','xi','phi'),
                                 verbose=FALSE,
                                 show_diagnostic_plots=FALSE,
                                 train_on_OE=FALSE)
estimates = readRDS(paste('fits/model_comparison',identifier,'.rds',sep='')) 
library(bayesplot)
#want to compare to draws from the prior for nu
nu_df = data.frame(posterior=as.numeric(as.array(estimates)[,,'nu']), prior=0.5 + 0.5*runif(4000)) %>% tidyr::gather(key=distribution,value=nu)
color_scheme_get("purple")
#g2 <- mcmc_dens(as.array(estimates),pars='nu') +
g2 <- ggplot(data=NULL) + geom_blank() + theme_bw() +
  theme(text = element_text(size = 12), axis.text = element_text(size = 12),
          strip.text = element_text(size = 8),legend.position=c(0.2,0.7)) + 
  labs(x=expression(paste('Transport bias  ', nu)),y='Density') + 
  geom_density(data=nu_df,aes(nu,fill=distribution,alpha=distribution)) + 
  scale_fill_manual(values = c('#e5cce5','#400040'),guide = guide_legend(title = 'Distribution')) + 
  scale_alpha_manual(values = c(0.66,0.66),guide = FALSE)
print(g2)
ggsave(paste('plots/fig4_nu_marginal',identifier,'.eps',sep=''),device=cairo_ps)
library(patchwork)
g1 <- ggplot(data=NULL) + geom_blank() + theme_minimal()
{g1 + g2} - res[[1]] + plot_layout(ncol=1)
ggsave(paste('plots/fig4',identifier,'.eps',sep=''),device=cairo_ps)
```

Fig5: Results of inference for dynamic model A) Posterior pairs plot, B) Posterior predictive distribution, C) Sensitivity to a, b.
```{r, Fig5, echo=FALSE,message=FALSE}
source('mrna_transport_full.R')
res_WT = mrna_transport_inference_full('no_decay_v331',
						use_real_data = TRUE, run_mcmc = FALSE,
						nSamples = 9, nTest = 11, nTestOE = 9,
						verbose = TRUE, compare_via_loo = FALSE,
						show_diagnostic_plots = TRUE, train_on_OE = FALSE,
						parametersToPlot = c('a','b','nu','phi','sigma'),alpha=2.18,model_str='simple')
res_OE = mrna_transport_inference_full('no_decay_OE_v332',
						use_real_data = TRUE, run_mcmc = FALSE,
						nSamples = 9, nTest = 11, nTestOE = 9,
						verbose = TRUE, compare_via_loo = FALSE,
						show_diagnostic_plots = TRUE, train_on_OE = TRUE,
						parametersToPlot = c('a','b','nu','phi','sigma'),
						alpha=2.18, model_str='simple')
source('figure5_plotting.R')
marginal_WT + res_WT + plot_layout(ncol=1,height=c(1,3))
```

![optional caption text](plots/posterior_predno_decay_v331.eps){ width=30% }

Fig7: Validation of testable prediction for over expression versus measured results.  
```{r, Fig7, echo=FALSE}
library(rstan)
library(broom)
library(dplyr)
identifier = 'fig7v001'
nSamples=9
nTest=11
nTestOE=9
source('extract_times_and_scaling.R')
times = extract_times_and_scaling(nSamples,nTest,nTestOE)
y0 = rep(1,16); y0[1]=0
stan_file = 'switch_off_RCs.stan'
stan_list = list(y0=y0,
                 T1  = nSamples,                 
                 t0 = times$t0$estimate[3],
                 ts1 = times$ts1,                 
                 a = 10,
                 b = 0.2,
                 gamma = 0,
                 nu = 0.9,
                 sigma = 1.25,
                 phi = 0.3,
                 b_i = 8,
                 b_j = 16)
estimates <- stan(file = stan_file,
                  data = stan_list,
                  seed = 42,
                  chains = 4,
                  warmup = 1000,
                  iter = 2000
)
tryCatch({
  #estimates@stanmodel@dso <- new("cxxdso") #seems have to do this to be able to save :S
  saveRDS(estimates, file = paste('fits/switch_off_RCs_',identifier,'.rds',sep='')) # to load use readRDS("fit.rds")
}, warning = function(war) {  
  # warning handler picks up where error was generated
  print(paste("WARNING:  ",war))
}, error = function(err) {  
  # error handler picks up where error was generated
  print(paste("NOT SAVED ERROR:  ",err))
}) # END tryCatch
source('compare_RC_preds.R')
comparisons <- compare_RC_preds(times$ts1,estimates,identifier)
comparisons[[1]]
```

Figure 9: to show change in number of complexes localized in the oocyte over time

```{r}
source('figure9.R')
```


FigS1: (Supplementary) Exponential growth model to establish timescales A) Schematic of time points used for each stage, B) Linear regression of log(A) against the time points for each stage.
```{r, FigS1, echo=FALSE}
source('extract_times_and_scaling.R')
extract_times_and_scaling(nSamples=9,nTest=11,nTestOE=9,optional_plot=TRUE)
```

FigS2: Convergence diagnostics for MCMC inference methods
![MCMC trace corresponding to Figure 4](plots/traceMCv321.eps){ width=30% }
![MCMC trace corresponding to Figure 5](plots/traceno_decay_v331.eps){ width=30% }


FigS3: Prior predictive distribution
```{r, FigS3, echo=FALSE}
source('mrna_transport_full.R')
res_PP = mrna_transport_inference_full('prior_pred_v391',
						use_real_data = TRUE, run_mcmc = TRUE,
						nSamples = 9, nTest = 11, nTestOE = 9,
						verbose = TRUE, compare_via_loo = FALSE,
						show_diagnostic_plots = FALSE, use_hierarchical_model = FALSE,
						use_prior_predictive = TRUE, train_on_OE = FALSE,
						is_nu_uniform = TRUE, no_decay_model = TRUE)
```

FigSPairs: Posterior pairwise plot
![optional caption text](plots/pairsno_decay_v331.eps){ width=30% }

FigS4: Sensitivity analysis
```{r, FigS4}
source('sensitivity_and_analytics.R')
```

