---
title: "Meeting with REB"
author: "Jonathan Harrison"
date: "07/09/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RNA localization
## Big picture

* Developed ODE model for RNA transport and fitted this full model to data to estimate parameters $(a, b, \gamma, \phi, \nu, \sigma)$
* Gained bioloical insight from parameter estimates
* Predictions from fitted model for wild type (WT) behaviour match well
* Predictions from fitted model for overexpression (OE) behaviour agree for nurse cells close to oocyte but disagree for more distant cells

## How can we change the model so that it agrees better for cells further from oocyte?
I have considered alternative models as follows:

* Spatially varying transport bias, $\nu$
* Spatially varying production in OE mutant
* Hierarchical model

Additional suggestion from Richard:
* Perturbations to entries of matrix $B$

## Details to discuss
### Timescales via exponential growth model
Need a precise age for each egg chamber. Can roughly estimate age via developmental time (eg. stage 7) by observing presence of distinctive features. Each stage is a large time window and would prefer a more precise age. By fitting an exponential growth model (independently shown to be a good description at these ages in other work, Jia et al 2016), we can use $\log(A)$, where $A$ is the area of a section through the egg chamber, as a proxy for time.

Do we need a separate exponential growth model for data from different experiments, or for different mutants?

### Assembly/aggregation parameter $\phi$
There is (at least one) structural unidentifiability in the full model (I think?).  

### Identifiability
Tried to assess identifiability of the full model using the symbolic computing capabilities in MATLAB. This involves solving matrix equations involving parameters ($\nu$ and $\gamma$), and was prohibitively slow. Could there be better ways to parameterise the model? Nondimensionalising allows us to remove a parameter, such as transport rate $b$, but to fit to data need to relate to dimensional quantities and dimensional time.  

### Notation for manuscript


### Switch off ring canals
Simulate from a model where we switch off selected ring canals and look at the effect of this on accumulation of RNA (eg. in cells far from the oocyte).
```{r}
nSamples=9
nTest=11
nTestOE=9
source('extract_times_and_scaling.R')
times = extract_times_and_scaling(nSamples,nTest,nTestOE)
y0 = rep(1,16); y0[1]=0
stan_file = 'switch_off_RCs.stan'
stan_list = list(y0=y0,
                 T1  = nSamples,                 
                 t0 = times$t0$estimate[3],
                 ts1 = times$ts1,                 
                 a = 9,
                 b = 0.2,
                 gamma = 0.01,
                 nu = 0.9,
                 sigma = 1.25,
                 phi = 0.57,
                 alter_matrix = 1,
                 b_i = 8,
                 b_j = 16)
estimates <- stan(file = stan_file,
                  data = stan_list,
                  seed = 42,
                  chains = 4,
                  warmup = 1000,
                  iter = 2000
)
```


